---
title: "Chimerism dataset final - bone marrow"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'plots_bmc.html'))})
author:
  - David C. Shyr^[Stanford Medicine, dcshyr@stanford.edu]
  - Simon Brewer^[University of Utah, simon.brewer@geog.utah.edu]
date: "03/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r message=FALSE}
library(tidyverse)
library(reshape2)
library(ggpubr)
library(lubridate)
```

## Data

```{r}
all.df = read.csv("./data/all.df.csv")
names(all.df)
all.df$ID = as.factor(all.df$ID)
```

Remove plot relapse samples
```{r}
plot.df = all.df[which(all.df$bdate < all.df$dor | is.na(all.df$dor)), ]
```

### CDW proportions

- Relapse vs. non-relapse
```{r warning=FALSE}
ggboxplot(plot.df, x = "relapse", y = "bmc_cdw", 
          color = "relapse", add = "jitter", main = "BM CD (Whole)") +
  stat_compare_means(label.y = 20, method = "wilcox.test") + 
  scale_y_continuous(limits = c(0,100))
```

- Time plots
```{r warning=FALSE}
ggscatter(plot.df, x = "btime", y = "bmc_cdw", col = "relapse", main = "BM CD (Whole)",
          add = "reg.line")
p1 = ggline(plot.df, x = "btime", y = "bmc_cdw", col = "ID", 
       facet.by = "relapse", numeric.x.axis = TRUE, main = "BM CD (Whole)")
ggpar(p1, legend = "none")
```

- Boxplots by testing period
```{r warning=FALSE}
ggboxplot(plot.df, x = "test", y = "bmc_cdw", fill = "relapse", main = "BM CD (Whole)")
```

### CD3 proportions

- Relapse vs. non-relapse
```{r warning=FALSE}
ggboxplot(plot.df, x = "relapse", y = "bmc_cd3", 
          color = "relapse", add = "jitter", main = "BM CD3") +
  stat_compare_means(label.y = 20, method = "wilcox.test") + 
  scale_y_continuous(limits = c(0,100))
```

- Time plots
```{r warning=FALSE}
ggscatter(plot.df, x = "btime", y = "bmc_cd3", col = "relapse", main = "BM CD3",
          add = "reg.line")
p1 = ggline(plot.df, x = "btime", y = "bmc_cd3", col = "ID", 
       facet.by = "relapse", numeric.x.axis = TRUE, main = "BM CD3")
ggpar(p1, legend = "none")
```

- Boxplots by testing period
```{r warning=FALSE}
ggboxplot(plot.df, x = "test", y = "bmc_cd3", fill = "relapse", main = "BM CD3")
```

### CD15 proportions

- Relapse vs. non-relapse
```{r warning=FALSE}
ggboxplot(plot.df, x = "relapse", y = "bmc_cd15", 
          color = "relapse", add = "jitter", main = "BM CD15") +
  stat_compare_means(label.y = 20, method = "wilcox.test") + 
  scale_y_continuous(limits = c(0,100))
```

- Time plots
```{r warning=FALSE}
ggscatter(plot.df, x = "btime", y = "bmc_cd15", col = "relapse", main = "BM CD15",
          add = "reg.line")
p1 = ggline(plot.df, x = "btime", y = "bmc_cd15", col = "ID", 
       facet.by = "relapse", numeric.x.axis = TRUE, main = "BM CD15")
ggpar(p1, legend = "none")
```

- Boxplots by testing period
```{r warning=FALSE}
ggboxplot(plot.df, x = "test", y = "bmc_cd15", fill = "relapse", main = "BM CD15")
```

### CD34 proportions

- Relapse vs. non-relapse
```{r warning=FALSE}
ggboxplot(plot.df, x = "relapse", y = "bmc_cd34", 
          color = "relapse", add = "jitter", main = "BM CD34") +
  stat_compare_means(label.y = 20, method = "wilcox.test") + 
  scale_y_continuous(limits = c(0,100))
```

- Time plots
```{r warning=FALSE}
ggscatter(plot.df, x = "btime", y = "bmc_cd34", col = "relapse", main = "BM CD34",
          add = "reg.line")
p1 = ggline(plot.df, x = "btime", y = "bmc_cd34", col = "ID", 
       facet.by = "relapse", numeric.x.axis = TRUE, main = "BM CD34")
ggpar(p1, legend = "none")
```

- Boxplots by testing period
```{r warning=FALSE}
ggboxplot(plot.df, x = "test", y = "bmc_cd34", fill = "relapse", main = "BM CD34")
```

## Comparative plots

```{r}
plot2.df = melt(plot.df, measure.vars = c("bmc_cdw", "bmc_cd3", "bmc_cd15", "bmc_cd34"),
                variable.name = "marker", value.name = "prop")
```

- Relapse vs. non-relapse
```{r warning=FALSE}
ggboxplot(plot2.df, x = "relapse", y = "prop", 
          fill = "marker", main = "BM All markers") +
  scale_y_continuous(limits = c(0,100))
```

- Boxplots by testing period
```{r warning=FALSE}
ggboxplot(plot2.df, x = "test", y = "prop", fill = "marker", 
          facet.by = "relapse", main = "BM All markers")
```
