---
title: "Chimerism dataset final - LIME plots"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'rf_lime.html'))})
author:
  - David C. Shyr^[Stanford Medicine, dcshyr@stanford.edu]
  - Simon Brewer^[University of Utah, simon.brewer@geog.utah.edu]
date: "3/24/2021"
output:
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Set random seed for reproducibility
```{r}
set.seed(1234)
```

```{r message=FALSE}
library(dplyr)
library(lubridate)
library(lime)       # ML local interpretation
library(vip)        # ML global interpretation
library(pdp)        # ML global interpretation
library(ggplot2)    # visualization pkg leveraged by above packages
library(caret)      # ML model building
```

## Data

Read in data
```{r}
all.df <- read.csv("./data/all.df.csv")
```

Convert dates
```{r}
all.df$dot <- ymd(all.df$dot)
all.df$dor <- ymd(all.df$dor)
all.df$bdate <- ymd(all.df$bdate)
all.df$pdate <- ymd(all.df$pdate)
```

Convert all character strings to factors
```{r}
all.df <- all.df %>% mutate_if(is.character,as.factor)
```

Make outcome a binary variable (0/1 relapse)
```{r}
all.df$rbin <- factor(all.df$rbin, levels = c("yes", "no"))
```

Filter out any tests that are post-relapse
```{r}
all.df <- all.df[which(all.df$bdate < all.df$dor | is.na(all.df$dor)), ]
```

Filter out relapse >720 days
```{r}
all.df <- all.df[which(all.df$rbin == "no" | all.df$rtime < 720),]
```

Filter out any missing tests
```{r}
all.df <- all.df[!is.na(all.df$bmc_cdw) & !is.na(all.df$bmc_cd3) & 
                   !is.na(all.df$bmc_cd15) & !is.na(all.df$bmc_cd34) &
                   !is.na(all.df$pbc_cdw) & !is.na(all.df$pbc_cd3) & 
                   !is.na(all.df$pbc_cd15) & !is.na(all.df$pbc_cd34),]
all.df <<- all.df
```

## Set up for LIME plots

Sub out the required data

rbin ~ txage + sex + rstatprtx + hla + 
tbi + gmgp + agvhd + cgvhd +
bmc_cdw + bmc_cd3 + bmc_cd15 + bmc_cd34 + 
pbc_cdw + pbc_cd3 + pbc_cd15 + pbc_cd34


```{r}
all.df2 <<- all.df %>%
  select(rbin, txage, sex, rstatprtx, hla, 
         tbi, gmgp, agvhd, cgvhd, ## Removed e as only 1 level
         bmc_cdw, bmc_cd3, bmc_cd15, bmc_cd34, 
         pbc_cdw, pbc_cd3, pbc_cd15, pbc_cd34)
```

### Set up random forest through caret

```{r}
# fit.caret <- train(
#   rbin ~ ., 
#   data = all.df2, 
#   method = 'rf',
#   trControl = trainControl(method = "cv", number = 5, classProbs = TRUE),
#   tuneLength = 1
# )

fit.caret <- train(
  rbin ~ ., 
  data = all.df2, 
  method = 'rf'
)
```

```{r}
fit.caret
```
Optional rf model --> probably not needed
```{r}
fit.rf <- randomForest::randomForest(
  rbin ~ ., 
  data = all.df2)
```


## variable importance plots

```{r}

vip(fit.rf, method = "permute", data = all.df2, target = "rbin", 
    metric = "auc", pred_wrapper = predict, 
    reference_class = "no", nsim = 10) + ggtitle("ranger: RF")

```

```{r}
# vis <- vi(fit.ranger, method = "permute", data = all.df2, target = "rbin", 
#           metric = "auc", pred_wrapper = pfun, 
#           reference_class = "no", nsim = 100) 
# vip(vis, geom = "boxplot") # Figure 12

vis <- vi(fit.rf, method = "permute", data = all.df2, target = "rbin", 
          metric = "auc", pred_wrapper = predict, 
          reference_class = "no", nsim = 100) 
vip(vis, geom = "boxplot") # Figure 12

```

```{r}
p <- ggplot(vis, aes(Variable, Importance))
p  + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) + 
  geom_errorbar(aes(ymin = Importance-StDev, 
                    ymax = Importance+StDev), width = 0.2) +
  coord_flip()

p <- ggplot(vis, aes(reorder(Variable, Importance), Importance)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) + 
  geom_errorbar(aes(ymin = Importance-StDev, 
                    ymax = Importance+StDev), width = 0.2) +
  coord_flip() + theme_bw() + scale_x_discrete(name = "Variable")
print(p)
#ggsave("./lime_plots/vip.pdf", plot = p)
```

## Explainers

```{r}
explainer_caret <- lime(all.df2, fit.caret, n_bins = 5)
# explainer_rf <- lime(all.df2, fit.rf, n_bins = 5)

summary(explainer_caret)
# summary(explainer_rf)
```

Example explainer plot for patient 1
```{r}
patientID <- which(all.df$ID == 1)

explanation_caret <- explain(
  x = all.df2[patientID,], 
  explainer = explainer_caret, 
  n_permutations = 5000,
  dist_fun = "gower",
  kernel_width = .75,
  n_features = 10, 
  feature_select = "highest_weights",
  labels = "yes"
)

p1 <- plot_features(explanation_caret)

plot_explanations(explanation_caret)
```

## All patients

```{r}
all_patients = unique(all.df$ID)
for (i in 1:length(all_patients)) {
  patientID <- which(all.df$ID == all_patients[i])
  
  explanation_caret <- explain(
    x = all.df2[patientID,], 
    explainer = explainer_caret, 
    n_permutations = 5000,
    dist_fun = "gower",
    kernel_width = .75,
    n_features = 10, 
    feature_select = "highest_weights",
    labels = "yes"
  )
  
  p1 <- plot_features(explanation_caret) + ggtitle(paste("Patient", all_patients[i]))
  print(p1)
  #ggsave(paste0("./lime_plots/patient_",i,".pdf"), plot = p1)
}
```


## Selected patients

Patient 8 (ID = 5)
```{r}
j = 5
patientID <- which(all.df$ID == j)

## Print one line of table to check
all.df[patientID[1], ]

for (i in patientID) {
  explanation_caret <- explain(
    x = all.df2[i,], 
    explainer = explainer_caret, 
    n_permutations = 5000,
    dist_fun = "gower",
    kernel_width = .75,
    n_features = 10, 
    feature_select = "highest_weights",
    labels = "yes"
  )
  p1 <- plot_features(explanation_caret) + ggtitle(paste("Patient", j))
  print(p1)
}
```

Patient 15 (ID = 43)
```{r}
j = 43
patientID <- which(all.df$ID == j)

## Print one line of table to check
all.df[patientID[1], ]

for (i in patientID) {
  explanation_caret <- explain(
    x = all.df2[i,], 
    explainer = explainer_caret, 
    n_permutations = 5000,
    dist_fun = "gower",
    kernel_width = .75,
    n_features = 10, 
    feature_select = "highest_weights",
    labels = "yes"
  )
  p1 <- plot_features(explanation_caret) + ggtitle(paste("Patient", j))
  print(p1)
}
```

Patient 16 (ID = 8)
```{r}
j = 8
patientID <- which(all.df$ID == j)

## Print one line of table to check
all.df[patientID[1], ]

for (i in patientID) {
  explanation_caret <- explain(
    x = all.df2[i,], 
    explainer = explainer_caret, 
    n_permutations = 5000,
    dist_fun = "gower",
    kernel_width = .75,
    n_features = 10, 
    feature_select = "highest_weights",
    labels = "yes"
  )
  p1 <- plot_features(explanation_caret) + ggtitle(paste("Patient", j))
  print(p1)
}
```


Patient 17 (ID = 45)
```{r}
j = 45
patientID <- which(all.df$ID == j)

## Print one line of table to check
all.df[patientID[1], ]

for (i in patientID) {
  explanation_caret <- explain(
    x = all.df2[i,], 
    explainer = explainer_caret, 
    n_permutations = 5000,
    dist_fun = "gower",
    kernel_width = .75,
    n_features = 10, 
    feature_select = "highest_weights",
    labels = "yes"
  )
  p1 <- plot_features(explanation_caret) + ggtitle(paste("Patient", j))
  print(p1)
}
```

