---
title: "Chimerism dataset final - Mixed Effects RF"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'rf_merf.html'))})
author:
  - David C. Shyr^[Stanford Medicine, dcshyr@stanford.edu]
  - Simon Brewer^[University of Utah, simon.brewer@geog.utah.edu]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Set random seed for reproducibility
```{r}
set.seed(1234)
```

```{r message=FALSE}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(randomForest)
library(vip)
library(pdp)
library(htree)
library(caret)
library(PresenceAbsence)
```

## Data

Read in data
```{r}
all.df <- read.csv("./data/all.df.csv")
```

Convert dates
```{r}
all.df$dot <- ymd(all.df$dot)
all.df$dor <- ymd(all.df$dor)
all.df$bdate <- ymd(all.df$bdate)
all.df$pdate <- ymd(all.df$pdate)
```

Convert all character strings to factors
```{r}
all.df <- all.df %>% mutate_if(is.character,as.factor)
```

Make outcome a binary variable (0/1 relapse)
```{r}
all.df$rbin <- factor(all.df$rbin, levels = c("yes", "no"))
```

Make a "time" variable
```{r}
all.df$time <- as.numeric(substr(all.df$test, 2, 2))
```

Filter out any tests that are post-relapse
```{r}
all.df <- all.df[which(all.df$bdate < all.df$dor | is.na(all.df$dor)), ]
```

Filter out relapse >720 days
```{r}
all.df <- all.df[which(all.df$rbin == "no" | all.df$rtime < 720),]
```

Filter out any missing tests
```{r}
all.df <- all.df[!is.na(all.df$bmc_cdw) & !is.na(all.df$bmc_cd3) & 
                   !is.na(all.df$bmc_cd15) & !is.na(all.df$bmc_cd34) &
                   !is.na(all.df$pbc_cdw) & !is.na(all.df$pbc_cd3) & 
                   !is.na(all.df$pbc_cd15) & !is.na(all.df$pbc_cd34),]
all.df <<- all.df
```

## Hierarchical random forest

First set the formula (note we need to drop `e` for this to work)
```{r}
f1 = rbin ~ txage + sex + rstatprtx + g + hla + 
  tbi + abd + ci + mtx + mmf + agvhd + cgvhd +
  bmc_cdw + bmc_cd3 + bmc_cd15 + bmc_cd34 + 
  pbc_cdw + pbc_cd3 + pbc_cd15 + pbc_cd34
```

## Cross validation

Next find patients for cross-validation

```{r}
patients = unique(sort(all.df$ID))
npatients = length(patients)
nfolds = 5
pfolds = createFolds(all.df$relapse, k = nfolds)
pfolds
```

Data frame for model results

```{r}
mod_results <- data.frame(folds = c(1:nfolds, "mean"),
                          auc = rep(NA, nfolds+1),
                          sens = rep(NA, nfolds+1),
                          spec = rep(NA, nfolds+1)
)
```

## $k$-fold loop

```{r}
for (i in 1:nfolds) {
  print(i)
  patient_id <- pfolds[[i]]
  training <- all.df[-patient_id,]
  testing <-  all.df[ patient_id,]
  
  ## Select predictors
  train_x <- training %>% 
    select(ID, time, txage, rbin, sex, rstatprtx, hla, 
           tbi, gmgp, agvhd, cgvhd, 
           bmc_cdw, bmc_cd3, bmc_cd15, bmc_cd34,
           pbc_cdw, pbc_cd3, pbc_cd15, pbc_cd34)
  
  test_x <- testing %>% 
    select(ID, time, txage, rbin, sex, rstatprtx, hla, 
           tbi, gmgp, agvhd, cgvhd, 
           bmc_cdw, bmc_cd3, bmc_cd15, bmc_cd34,
           pbc_cdw, pbc_cd3, pbc_cd15, pbc_cd34)
  
  control<-list(se = TRUE, ntrees = 500, R = 100)
  
  # ff<-hrf(x = train_x, id = training$ID,
  #        time = training$time, yindx = "rbin",
  #        mtry = 5,
  #        historical = TRUE,
  #        control = control)
  ff <- hrf(x = train_x, id = train_x$ID,
            time = train_x$time, yindx = "rbin",
            ntrees = 500, 
            mtry = 5,
            historical = TRUE)
  
  pred_test <- predict_hrf(ff, test_x)
  
  df <- data.frame(id = seq(1:nrow(testing)),
                   obs = as.numeric(testing$rbin)-1,
                   pred = pred_test[,1])
  
  mod_results$auc[i] <- auc(df)$AUC[[1]]
  ## Get threshold
  opt_thresh <- optimal.thresholds(df, opt.methods = 3)[2]
  opt_cmx <- cmx(df, threshold = opt_thresh$pred)
  mod_results$sens[i] = sensitivity(opt_cmx)$sensitivity[[1]]
  mod_results$spec[i] = specificity(opt_cmx)$specificity[[1]]
  
}
```

```{r}
mod_results[6, 2:4] = apply(mod_results[, 2:4], 2, mean, na.rm = TRUE)
knitr::kable(mod_results, digits = 4)
```
